---
title: "Store Analysis"
author: "Brendan Callender"
format: html
editor: visual
---

## Setup

```{r}
library(effects)
library(tidyverse)
library(pivottabler)
library(car)
library(performance)
library(fitdistrplus) #fitdist
library(MASS) #glm.nb
library(pscl) #odTest, zeroinfl, hurdle
```

### Read in Data

```{r}
here::here()
Store <- read_csv(here::here("portfolio", "glm", "Store.csv"))
Store$logunits <- log(Store$units)
```

## Initial Model

```{r}
m1 <- glm(customers ~ income + age + compdist + storedist, family = poisson(link = "log"), offset = logunits, data = Store)
summary(m1)
```

### Goodness of Fit

Strong evidence that poisson loglinear model with all main effects does not adequately fit the data...

```{r}
1- pchisq(462.34, 110-5) # deviance gof
1- pchisq(sum(residuals(m1, type = "pearson")^2), 110-5) # pearson gof
```

### Testing Overdispersion

We have strong evidence that the poisson loglinear model with all main effects has overdispersion so we will have to find a solution for that

```{r}
check_overdispersion(m1)
```

## Addressing Overdispersion

### NB2 Model

```{r}
m2 <- glm.nb(customers ~ income + age + compdist + storedist + offset(logunits), data = Store)
summary(m2)
```

### QL (Var = mu)

```{r}
m3 <- glm(customers ~ income + age + compdist + storedist, family = quasi(link = "log", var = "mu"), offset = logunits, data = Store)
summary(m3)
```

### QL (Var = mu\^2)

```{r}
m4 <- glm(customers ~ income + age + compdist + storedist, family = quasi(link = "log", var = "mu^2"), offset = logunits, data = Store)
summary(m4)
```

### Comparison of Model $X^2$ Statistics

```{r}
x_squared <- c(sum(residuals(m1, type = "pearson")^2),
  sum(residuals(m2, type = "pearson")^2),
  sum(residuals(m3, type = "pearson")^2),
  sum(residuals(m4, type = "pearson")^2))
  
names(x_squared) <- c("Poisson", "NegBin", "Quasi(mu)", "Quasi(mu^2)")
dfs <- c(105,104,105,105)
x_squared
pchisq(x_squared, dfs, lower.tail = FALSE)
```

## Stepwise Model Selection for NB2 Model

```{r}
step.model.nb <- step(m2, scope = ~ .^3, k = 2, direction = "both", trace = 0)
summary(step.model.nb)
sum(residuals(step.model.nb, type = "pearson")^2)
```

```{r}
ts1 <- sum(residuals(step.model.nb, type = "pearson")^2)
ts2 <- sum(residuals(step.model.nb, type = "deviance")^2)
df <- 104
pchisq(ts1, df, lower.tail = FALSE)
pchisq(ts2, df, lower.tail = FALSE)
```

### Examine Diagnostic Plots for Stepwise NB2 Model

```{r}
car::infIndexPlot(step.model.nb, vars=c("Studentized", "hat", "Cook"))
residualPlots(step.model.nb, tests = FALSE)
```

### Examine Major Outlier

```{r}
Store[which.max(residuals(step.model.nb, type = "pearson")),]
```

## Redoing Analysis w/ Outlier Removed

```{r}
Store_sub <- Store[-11,]

m1_sub <- glm(customers ~ income + age + compdist + storedist, family = poisson(link = "log"), offset = logunits, data = Store_sub)

m2_sub <- glm.nb(customers ~ income + age + compdist + storedist + offset(logunits), link = "log", data = Store_sub)

m3_sub <- glm(customers ~ income + age + compdist + storedist, family = quasi(link = "log", var = "mu"), offset = logunits, data = Store_sub)

m4_sub <- glm(customers ~ income + age + compdist + storedist, family = quasi(link = "log", var = "mu^2"), offset = logunits, data = Store_sub)

```

### Comparing $X^2$ statistics with Outlier Removed

```{r}
x_squared <- c(
    sum(residuals(m1_sub, type = "pearson")^2),
    sum(residuals(m2_sub, type = "pearson")^2),
    sum(residuals(m3_sub, type = "pearson")^2),
    sum(residuals(m4_sub, type = "pearson")^2)
  )
names(x_squared) <- c("Poisson", "NegBin", "Quasi(mu)", "Quasi(mu^2)")
dfs <- c(105,104,105,105)
x_squared
pchisq(x_squared, dfs, lower.tail = FALSE)
```

## Examining QL(Var = mu\^2) Model

### QL(var = mu\^2)

Looking at the plot of the residuals, it looks like this model may have underdispersion and is overestimating the variance in the data since almost all preds are within 1 sd of the actual value and all others fall within 2 sds

```{r}
summary(m4_sub)
```

```{r}
infIndexPlot(m4_sub, vars = c("Studentized", "hat", "cook"))
```

```{r}
#| warning: false
residualPlot(m4_sub)
# residualPlots(m4_sub, tests = FALSE)
```

## Continuing Analysis with NB2 Model

```{r}
step.model.nb.sub <- step(m2_sub, scope = ~ .^3, k = 2, direction = "both", trace = 0)
summary(step.model.nb.sub)
sum(residuals(step.model.nb.sub, type = "pearson")^2)
```

### Model VIFs

```{r}
#| warning: false
#| message: false
vif(step.model.nb.sub)
```

### Centering Predictors

```{r}
Store_sub <- Store_sub %>%
  mutate(
    cage = scale(age, center = TRUE, scale = TRUE)[,1],  
    cincome = scale(income, center = TRUE, scale = TRUE)[,1],   
    cstoredist = scale(storedist, center = TRUE, scale = TRUE)[,1],
    ccompdist = scale(compdist, center = TRUE, scale = TRUE)[,1]
  )
```

```{r}
nb.main.cent.sub <- glm.nb(customers ~ cage + cincome + cstoredist + ccompdist + offset(logunits), link = log, data = Store_sub)

step.nb.cent.sub <- step(nb.main.cent.sub, scope = ~ .^3, k = 2, direction = "both", trace = 0)
summary(step.nb.cent.sub)
sum(residuals(step.nb.cent.sub, type = "pearson")^2)
```

### Check VIFs again

```{r}
#| warning: false
#| message: false
vif(step.nb.cent.sub)
```

## Final Model Hypothesis Testing

## Final Model Hypothesis Testing

```{r}
final_model <- step.nb.cent.sub
summary(final_model)
```

### Remove Non-Significant Interaction

```{r}
final_model <- update(final_model, ~ . - cstoredist:ccompdist)
summary(final_model)
```

### Add Age?

```{r}
m.add.age <- update(final_model, ~ . + age)
summary(m.add.age)
```

### Add Any Interactions?

```{r}
all.int <- glm.nb(customers ~ (income + age + compdist + storedist)^2 + offset(logunits), data = Store_sub)
anova(final_model, all.int)
```

### Add Higher Order Terms?

```{r}
all.hot <- glm.nb(customers ~ poly(cincome, 2) + poly(cage,2) + poly(ccompdist,2) + poly(cstoredist, 2) + offset(logunits), data = Store_sub)
anova(final_model, all.hot)
```

Higher Order term for Income is close to being a significant predictor at a 5% overall significance level.

```{r}
summary(glm.nb(customers ~ poly(cincome, 2) + ccompdist + cstoredist + offset(logunits), data = Store_sub))
```

## Final Model Diagnostics

```{r}
summary(final_model)
```

```{r}
vif(final_model)
```

### Goodness of Fit Test

```{r}
ts <- sum(residuals(final_model, type = "pearson")^2)
pchisq(ts, 104, lower.tail = FALSE)
# df = 109 - 5 (p includes estimate for gamma)
```

Check Expected Counts / fitted values to make sure Goodness of fit test is valid GOF test is good since no expected counts are below 1 and only 9% of counts are below 5.

```{r}
exp_counts <- final_model$fitted.values

mean(exp_counts < 1)
mean(exp_counts < 5)
```

### Leverage / Influence

```{r}
n <- 109
p <- 5
high_leverage_cutoff <- 3*p/n
high_influence_cutoff <- qf(0.5, p, n-p)

c(high_leverage_cutoff, high_influence_cutoff)
```

There are no highly influential observations in the data

```{r}
infIndexPlot(final_model, vars = c("hat", "cook"))
resid_plots <- residualPlots(final_model, smooth = FALSE, tests = FALSE)
attr(resid_plots, "dimnames")[[1]] <- c("Income (C)", "Distance to Store (C)", "Distance to Competitor (C)")
resid_plots
```

### Examine Obs with high leverage

These observations appear to be neighborhoods with high median income. Overall, these points are not influential so they will be kept in the model.

```{r}
Store_sub[c(14, 93),]
```

## Outputting Final Model Predictions

```{r}
units = Store_sub %>%
  summarize(mean = mean(units)) %>%
  unlist() %>% unname() %>% data.frame(units = .) %>% round(., digits = 0)

cincome = Store_sub %>%
  summarize(min = min(cincome), mean = mean(cincome), max = max(cincome)) %>%
  unlist() %>% unname() %>% data.frame(cincome = .) %>% round(., digits = 2)

ccompdist = Store_sub %>%
  summarize(min = min(ccompdist), mean = mean(ccompdist), max = max(ccompdist)) %>%
  unlist() %>% unname() %>% data.frame(ccompdist = .) %>% round(., digits = 2)

cstoredist = Store_sub %>%
  summarize(min = min(cstoredist), mean = mean(cstoredist), max = max(cstoredist)) %>%
  unlist() %>% unname() %>% data.frame(cstoredist = .) %>% round(., digits = 2)

pred_data <- crossing(units, cincome, ccompdist, cstoredist) %>%
  mutate(logunits = log(units))
```

```{r}
pred_data %>%
  mutate(pred_count = predict(final_model, newdata = pred_data, type = "response")) %>%
  dplyr::select(-logunits) %>%
  mutate(pred_count = round(pred_count, digits = 3)) %>%
  arrange(desc(pred_count)) 
# %>%
#   knitr::kable(format = "html")
```
